---
title: "Developer Notes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What does it do?

`lexl` tokenises Excel formulas into some kind of parse tree, represented by a
data frame with three columns: the tokens themselves, the type,
and the level within the nested structure.

## How does tokenise formulas?

The bulk of the work is done by a parser generator, according to a grammer
defined in `src/token_grammar.h`.

### What is a parser generator?

A parser generator is a kind of supercharged REGEX engine, which allows actions
to be carried out upon each successful match of a particular group within the
overall pattern.  I don't know much about parser generators, so the rest of this

For example, in REGEX you might define a group by enclosing it within
parentheses, so that `Na(na)* Batman` would match `Nanana Batman` as well as
`Nananananananana Batman`.  With a parser generator, the pattern between the
parentheses is called a 'rule', and there is nothing that isn't a rule -- even
string literals must be set up as rules.

Any action defined for a given rule/group/pattern/what-you-will is carried out
whenever the rule is successfully matched.  For example, you could define an
action that prints `"Hey Jude?"`, to be carried out when the rule `(na)` is
matched.  Given the Batman string above, it that action would be performed
several times.

Rules can be disabled, and there are some tricks you can pull to have them act
differently depending on the current state, but really they're pretty
inflexible.  The most annoying problem is backtracking.

### Backtracking

This can happen when a rule is composed of subrules.  The parser attempts to
match the whole rule, and does not fail until it has already matched some of the
subrules.  The parser will 'wind the tape back' (go back to the place in the
string where it started to match the overall rule), but any actions defined on
the subrules will already have been performed.

For example, if the overall rule is `Na(na)* Batman`, and the string to be
matched is `Nananananananana Robin`, then `"Hey Jude?"` would still be printed
several times before the overall rule failed to match.

`xlex` avoids this problem by disabling certain rules.  The trade-off is that it
the resulting parse tree isn't very detailed.  If it needs to become more
detailed, then I must either become wickedly clever, or re-implement the grammer
using `boost::spirit` (which can do something about backtracking, albeit slow),
or parse the affected tokens a second time, using another grammer.

## TODO

### Pretty print abstract syntax tree in console

A minimum thing would be like:

```
print(lex_xl("MIN(3,MAX(2,A1))"))
\- root
 \- MIN
 \- (
  \- 3
  \- ,
  \- MAX
  \- (
   \- 2
   \- ,
   \- A1
  \- )
 \- )
```

Or without the parentheses

```
print(lex_xl("MIN(3,MAX(2,A1))"))
\- root
 \- MIN
  \- 3
  \- ,
  \- MAX
   \- 2
   \- ,
   \- A1
```

Or in pseudo-`plyr::ast()` [style](http://adv-r.had.co.nz/Expressions.html) (via
Jenny Bryan), thinking of something clever for operators.  Excel doesn't have
inline function definitions, and function arguments are positional, not named,
so pairlists can be ignored.

```
print(lex_xl("MIN(3,MAX(2,A1))"))
\- root
 \- ()
  \- `MIN
  \- 3
  \- ()
   \- `MAX
   \- 2
   \- `A1
```

Operators would have to be handled using associativity.  Fortunately, Excel
operators are consistently left-associative (so in Excel `2^2^3` is 64, whereas
in R and most other languages it is 256), which makes the logic easier.

#### API

* As an S3 override of `plot()`?
* As a separate function `lex_tree()`?

I think it's important that users realise it's just a data frame.  How does
data.tree do it?

### Evaluate formulas

This might well be possible using `shiny`.  Joe Cheng kindly wrote me an
demo of general reactivity.

```{r, eval = FALSE}
library(shiny)

a <- reactiveVal(0)
b <- reactive(-a())
c <- observe(print(b()))

shiny:::flushReact()


workbook <- new.env()

# For each formula cell
workbook[[cell_id]] <- reactive({
  # formula
  workbook[["other_cell"]]()
})

# For each value cell
workbook[[cell_id]] <- reactiveVal(initialValue)
```

References to external files (`[0]Sheet1!A1`) would have to be resolved to find
their value.  Fortunately these values are written into the local file.

### Improve coverage

Does RSelenium work on remote virtual machines?  Only if they have a real
browser installed.

https://cran.r-project.org/web/packages/RSelenium/vignettes/shinytesting.html,
local copy via `vignette("shinytesting", "RSelenium")`.
